---
title: "Recession Analysis and Visualizations"
output: flexdashboard::flex_dashboard
---

```{r Libraries, include=FALSE, message=FALSE}
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse) ; theme_set(theme_classic())

if (!require("DT")) install.packages("DT")
library(DT)

if (!require("Hmisc")) install.packages("Hmisc")
library(Hmisc)

if (!require("lubridate")) install.packages("lubridate")
library(lubridate)

if (!require("plotly")) install.packages("plotly")
library(plotly)

if (!require("gridExtra")) install.packages("gridExtra")
library(gridExtra)

if (!require("wesanderson")) install.packages("wesanderson")
library(wesanderson)

if (!require("viridis")) install.packages("viridis")
library(viridis)

if (!require("maps")) install.packages("maps")
library(maps)

if (!require("ggthemes")) install.packages("ggthemes")
library(ggthemes)

if (!require("shiny")) install.packages("shiny")
library(shiny)

if (!require("shinydashboard")) install.packages("shinydashboard")
library(shinydashboard)


if (!require("scales")) install.packages("scales")
library(scales)
```

```{r Global Setup, include=FALSE, echo=FALSE, message=FALSE}

# Setting up some global options for the knitr.
# General principle is that local options over-ride the global.

knitr::opts_chunk$set(
	#Chunk Code Options
	echo = FALSE, # Show the code in the output
	include = TRUE, # Show the results in the output.
	results = 'asis', 
	comment = NA, 
	prompt = FALSE,
	
	#Display Options
	cache = FALSE,

	# Plot Options
	fig.width = 7,
	fig.height = 5,
	
	fig.retina = 2,
	fig.align = 'center'
)

options(digits = 4)


```

```{r Load Data, include=FALSE, echo=FALSE, message=FALSE}
#We can use the load data command to load the data from the RData file we saved.

load("~/UNCG/IAF 605/R Assignment 1/Data/RecessionData.RData")

# And get its summary.
summary(RecessionData)
```

```{r Clean Data, include = FALSE, echo = FALSE, message = FALSE}
#Let's start by focusing in on the states of interest

CAMIRecessionData <- RecessionData %>% 
	filter((RecessionData$state == "California") | (RecessionData$state == "Massachusetts") | 
		   	(RecessionData$state == "Florida") |								(RecessionData$state == "North Carolina"))

# Remove the factors by converting to characters
CAMIRecessionData$state <- CAMIRecessionData$state %>% as.character()
str(CAMIRecessionData)

# Re-factor to two Factos - California and Michigan
CAMIRecessionData$state <- CAMIRecessionData$state %>% as.factor()

```

Where and When?
=====================================
Column {data-width=600}
-------------------------------------

### New Private Housing Contracts

```{r}
#Now let's start by creating the plots from the first Tableau Dashboard

#Let's start with an Area Plot of New Private Housing Contracts

newp_area <- ggplot(CAMIRecessionData, aes(x=date,
	   	y = new_private_housing_structures_authorized_by_building_permits_units, fill = state)) +
	geom_area(alpha = 0.6, na.rm = T) +
	scale_fill_viridis(discrete = T) +
	labs(title = "New Private Housing Contracts",
		 x = "Year",
		 y = "New Private Housing Structures Authorized by
		 Building Permits (Units)",
		 fill = "State")

ggplotly(newp_area, tootltip='text', originalData = F)

```

```{r CALI_H, include = FALSE, echo = FALSE, message = FALSE}
#California

HousingData <- RecessionData %>%
	filter(state == 'California') %>%
	group_by(year) %>%
	summarise(new_housing = sum(new_private_housing_structures_authorized_by_building_permits_units, na.rm = T)) %>%
	mutate(pct_change = (new_housing/lag(new_housing) - 1) * 100) %>%
	filter(!year >= 2020) %>%
	drop_na()


pc_cali <- ggplot(HousingData, aes(x = year, y = pct_change)) +
	geom_bar(stat = 'identity', aes(fill=pct_change)) +
	scale_fill_gradient2_tableau('pct_change', palette = "Orange-Blue Diverging") +
	coord_flip() +
	labs(title = "California",
		 x = "Year",
		 y = "% Change in New Private Housing Contracts",
		 		 fill = "Percent Change") +
	theme(legend.position = 'none')

pc_cali
```

```{r Fl_H, include = FALSE, echo = FALSE, message = FALSE}
#Florida

HousingData <- RecessionData %>%
	filter(state == 'Florida') %>%
	group_by(year) %>%
	summarise(new_housing = sum(new_private_housing_structures_authorized_by_building_permits_units, na.rm = T)) %>%
	mutate(pct_change = (new_housing/lag(new_housing) - 1) * 100) %>%
	filter(!year >= 2020) %>%
	drop_na()


pc_flor <- ggplot(HousingData, aes(x = year, y = pct_change)) +
	geom_bar(stat = 'identity', aes(fill=pct_change)) +
	scale_fill_gradient2_tableau('pct_change', palette = "Orange-Blue Diverging") +
	coord_flip() +
	labs(title = "Florida",
		 x = "Year",
		 y = "% Change in New Private Housing Contracts",
		 		 fill = "Percent Change") +
	theme(legend.position = 'none')

pc_flor
```

```{r mass_H, include = FALSE, echo = FALSE, message = FALSE}
#Mass

HousingData <- RecessionData %>%
	filter(state == 'Massachusetts') %>%
	group_by(year) %>%
	summarise(new_housing = sum(new_private_housing_structures_authorized_by_building_permits_units, na.rm = T)) %>%
	mutate(pct_change = (new_housing/lag(new_housing) - 1) * 100) %>%
	filter(!year >= 2020) %>%
	drop_na()

pc_mass <- ggplot(HousingData, aes(x = year, y = pct_change)) +
	geom_bar(stat = 'identity', aes(fill=pct_change)) +
	scale_fill_gradient2_tableau('pct_change', palette = "Orange-Blue Diverging") +
	coord_flip() +
	labs(title = "Massachusetts",
		 x = "Year",
		 y = "% Change in New Private Housing Contracts",
		 		 fill = "Percent Change") +
	theme(legend.position = 'none')

pc_mass

```

```{r nc_H, include = FALSE, echo = FALSE, message = FALSE}
#NC

HousingData <- RecessionData %>%
	filter(state == 'North Carolina') %>%
	group_by(year) %>%
	summarise(new_housing = sum(new_private_housing_structures_authorized_by_building_permits_units, na.rm = T)) %>%
	mutate(pct_change = (new_housing/lag(new_housing) - 1) * 100) %>%
	filter(!year >= 2020) %>%
	drop_na()


pc_nc <- ggplot(HousingData, aes(x = year, y = pct_change)) +
	geom_bar(stat = 'identity', aes(fill=pct_change)) +
	scale_fill_gradient2_tableau('pct_change', palette = "Orange-Blue Diverging") +
	coord_flip() +
	labs(title = "North Carolina",
		 x = "Year",
		 y = "% Change in New Private Housing Contracts",
		 		 fill = "Percent Change") +
	theme(legend.position = 'none')

pc_nc

```

### New Private Housing Contracts (% Change)
```{r}
#Grouping them together
grid.arrange(pc_cali, pc_flor, pc_nc, pc_mass, nrow = 2)

```

Column {data-width=400}
-------------------------------------
### New Private Housing Contracts (%Change All)
```{r}
new_house <- RecessionData %>%
	group_by(state, year) %>%
	arrange(year) %>%
	summarise(new_houses = sum(new_private_housing_structures_authorized_by_building_permits_units, na.rm = T)) %>%
	mutate(pct_change = (new_houses/lag(new_houses) - 1) * 100) %>%
	filter(!year >= 2020) %>%
	drop_na() %>%
	rename(State = state, Year = year, YoY_Change = pct_change)


new_house %>% 
  arrange(YoY_Change) %>% 
  head(1000) %>% 
  select(State, Year, YoY_Change) %>% 
  DT::datatable(options = list(pageLength=100)) %>%
  formatPercentage(columns = 'YoY_Change') %>% 
  formatStyle(.,    'YoY_Change',
    background = styleColorBar(new_house$YoY_Change, 'steelblue'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )
```

Home Vacancy Map {data-orientation=rows}
=====================================

Row {data-height=600}
-------------------------------------
### Recession
```{r}
#Let's Move on to a map of Home Vacancy Rates

if (!require("USAboundaries")) install.packages("USABoundaries")
library(USAboundaries)
if (!require("sf")) install.packages("sf")
library(sf)

#setup base layer
states <- us_states()
states <- states %>%
	filter(!name %in% c('Alaska', 'Hawaii', 'Puerto Rico', 'District of Columbia'))

HousingData <- RecessionData %>%
	filter(period == "Recession") %>%
	group_by(state) %>%
	summarise(MeanVacancy = mean(home_vacancy_rate_percent, na.rm = T))

HousingData$name <- HousingData$state

merged_data <-left_join(states, HousingData[, c('name', 'MeanVacancy')], by='name')

vacancy_du <- ggplot(merged_data) +
	geom_sf(aes(fill = MeanVacancy)) +
		scale_fill_gradient2_tableau('MeanVacancy', palette = "Orange-Blue Diverging") +
	labs(title = "Mean Vacancy Rate (2007-2009)")

ggplotly(vacancy_du)
```
Row {data-height=400}
-------------------------------------
### Pre-Recession
```{r}
HousingData <- RecessionData %>%
	filter(period == "Pre-Recession") %>%
	group_by(state) %>%
	summarise(MeanVacancy = mean(home_vacancy_rate_percent, na.rm = T))

HousingData$name <- HousingData$state

merged_data <-left_join(states, HousingData[, c('name', 'MeanVacancy')], by='name')

vacancy_pre <- ggplot(merged_data) +
	geom_sf(aes(fill = MeanVacancy)) +
		scale_fill_gradient2_tableau('MeanVacancy', palette = "Orange-Blue Diverging") +
	labs(title = "Mean Vacancy Rate (2000-2007)")

ggplotly(vacancy_pre)
```

### Post-Recession
```{r}
HousingData <- RecessionData %>%
	filter(period == "Post-Recession") %>%
	group_by(state) %>%
	summarise(MeanVacancy = mean(home_vacancy_rate_percent, na.rm = T))

HousingData$name <- HousingData$state

merged_data <-left_join(states, HousingData[, c('name', 'MeanVacancy')], by='name')

vacancy_po <- ggplot(merged_data) +
	geom_sf(aes(fill = MeanVacancy)) +
		scale_fill_gradient2_tableau('MeanVacancy', palette = "Orange-Blue Diverging") +
	labs(title = "Mean Vacancy Rate (2010-2017)")
	
ggplotly(vacancy_po)
```

Who?
=====================================
### Home Vacancy
```{r}
home_va <- ggplot(CAMIRecessionData, aes(x = year, y = home_vacancy_rate_percent)) +
	geom_smooth(aes(color=state), se = F, na.rm = T) +
	scale_color_viridis(discrete = T) +
	labs(title = "Home Vacancy",
		 x = "Year",
		 y = "Home Vacancy Rate (%)",
		 color = "State")
home_va
```

### Home Ownership
```{r}
home_ow <- ggplot(CAMIRecessionData, aes(x = year, y = home_ownership_rate_percent)) +
	geom_smooth(aes(color=state), se = F, na.rm = T) +
	scale_color_viridis(discrete = T) +
	labs(title = "Home Ownership",
		 x = "Year",
		 y = "Home Ownership Rate (%)",
		 color = "State")
home_ow
```

```{r Grouped Grid, include = FALSE, echo = FALSE, message = FALSE}
#Grouping it together

grid.arrange(home_va,home_ow, ncol = 1)
```

Waitress Index
=====================================

### Employees
```{r}

#Filter the years
restaurant <- CAMIRecessionData %>%
	filter( year < 2016 & year > 2005)


ggplot(restaurant, aes(x=year, fill = state)) +
	geom_bar(stat = 'identity', aes(y = all_employees_leisure_and_hospitality_food_service_and_drinking_places_in_thousands), na.rm = T) +
	scale_fill_viridis(discrete = T) +
	facet_grid(~state) +
		labs(title = "Employees Food Service",
		 x = "Year",
		 y = "Number Employees (thousands)") +
	theme(legend.position = 'none',
		  axis.text.x = element_text(angle = 90))
```

### Earnings
```{r}

#Calculate the mean
restaurant <- CAMIRecessionData %>%
	filter( year < 2016 & year > 2005) %>%
	group_by(state, year) %>%
	summarise(sum_food = sum(accomodation_and_food_service_earnings_in_thousands, na.rm = T), sum_employees = sum(all_employees_leisure_and_hospitality_food_service_and_drinking_places_in_thousands, na.rm = T), food_earnings = sum_food/sum_employees)

ggplot(restaurant, aes(x=year, fill = state)) +
	geom_bar(stat = 'identity', aes(y = food_earnings), na.rm = T) +
	scale_fill_viridis(discrete = T) +
	scale_y_continuous(labels = dollar) +
	facet_grid(~state) +
		labs(title = " Food Service Earnings",
		 x = "Year",
		 y = "Average Annual Earnigns Per Worker") +
	theme(legend.position = 'none',
		  axis.text.x = element_text(angle = 90))
```

Education {data-orientation=rows}
=====================================

Row {data-height=600}
-------------------------------------
### Graduate or Professional
```{r}
#Grad Degree
ggplot(CAMIRecessionData, aes(x=year, color = state)) +
	geom_smooth(aes(y = people_25_and_older_who_have_competed_a_graduate_or_professional_degree_percent), na.rm = T, se = F) +
	scale_color_viridis(discrete = T) +
		labs(title = "Percent of People 25 and older with a Graduate or Professional Degree",
		 x = "Year",
		 y = "Percent")
```

Row {data-height=400}
-------------------------------------
### Bachelor
```{r}
#Bachelor's
ggplot(CAMIRecessionData, aes(x=year, color = state)) +
	geom_smooth(aes(y = bachelors_degree_or_higher_percent), na.rm = T, se = F) +
	scale_color_viridis(discrete = T) +
		labs(title = "Percent of People with a Bachelor's Degree (or Higher)",
		 x = "Year",
		 y = "Percent")
```

### High School
```{r}
#High School
ggplot(CAMIRecessionData, aes(x=year, color = state)) +
	geom_smooth(aes(y = high_school_graduate_or_higher_percent), na.rm = T, se = F) +
	scale_color_viridis(discrete = T) +
		labs(title = "Percent of People that Graduated High School",
		 x = "Year",
		 y = "Percent")
```

Entrepreneurship
=====================================

Row {data-height=600}
-------------------------------------

### Business Formations
```{r}
#Business Formations
ggplot(CAMIRecessionData, aes(x=year, color = state)) +
	geom_smooth(aes(y = business_formation_within_four_quarters), na.rm = T, se = F) +
	scale_color_viridis(discrete = T) +
		labs(title = "Business Formations within 4 Quarters",
		 x = "Year",
		 y = "Business Formations",
		 fill = "State")
	
```

Row {data-height=400}
-------------------------------------
### Business Applications
```{r}
#NC / MA
NCMARecessionData <- CAMIRecessionData %>%
	filter(state == 'North Carolina' |
		   	state == 'Massachusetts')


ggplot(NCMARecessionData, aes(x=date,
	   	y = business_applications, fill = state)) +
	stat_smooth(geom = 'area', alpha = 0.6, na.rm = T, span = 1/4) +
	scale_fill_manual(values = c("#35B779FF", "#FDE725FF")) +
	facet_grid(~state)+
	labs(title = "Business Applications",
		 x = "Year",
		 y = "Business Applications",
		 fill = "State")
```

### Business Applications
```{r}
#CA / FL
CAFLRecessionData <- CAMIRecessionData %>%
	filter(state == 'California' |
		   	state == 'Florida')

ggplot(CAFLRecessionData, aes(x=date,
	   	y = business_applications, fill = state)) +
	stat_smooth(geom = 'area', alpha = 0.6, na.rm = T, span = 1/4) +
	scale_fill_manual(values = c("#440154FF", "#31688EFF")) +
	facet_grid(~state)+
	labs(title = "Business Applications",
		 x = "Year",
		 y = "Business Applications",
		 fill = "State")

```
